---
title: "Plotly_Assignment"
author: "Wanjie Feng"
date: "2024-04-04"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

In this assignment, a presentation using Plotly was created. 



## What is Plotly?

Plotly is a web application for creating and sharing data visualizations. Plotly
can work with several programming languages and applications including R,
Python, and Microsoft Excel. We're going to concentrate on creating different
graphs with Plotly and sharing those graphs.


## Summary of MTCARS

```{r}
summary(mtcars)
```

## Plot weight vs mpg

```{r , echo=FALSE, message=FALSE}
library(plotly)
library(ggplot2)
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter")
```

## Plot cylinder vs mpg

```{r , echo = FALSE, message=FALSE}
library(plotly)
plot_ly(mtcars, x = ~cyl, y = ~mpg, type = "box")
```

## Plot weight vs mpg using cylinder as factor and horsepower as size

```{r , echo = FALSE, message=FALSE, warning = FALSE}
library(plotly)
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", 
        color = ~factor(cyl), size = ~hp)
```

## Thank you. 